refnx -- Neutron and X-ray reflectometry analysis in Python

### 2018-07-31 paper/manuscript.tex
Allow hyphenation of further words

@@

\begin{abstract}
\refnx\ is a model-based neutron and X-ray reflectometry data analysis package written in Python. It is cross platform, and has been tested on Linux, macOS, and Windows. Its graphical user interface is browser-based, through a \Jupyter\ notebook.
Model construction is modular, being composed from a series of components that each describe a subset of the interface, parameterised in terms of physically relevant parameters (volume fraction of a polymer, lipid area per molecule, etc). The model and data are used to create an objective, which is used to calculate residuals, log-likelihood, and log-prior probabilities of the system. Objectives are combined to perform co-refinement of multiple datasets, and mixed-area models. Prior knowledge of parameter values is encoded as probability distribution functions or bounds on all parameters in the system. Extra prior probability terms can be defined for sets of components, over and above those available from the parameters alone. Algebraic parameter constraints are available.
A choice of fitting approaches is available, including least-squares (global and gradient-based optimizers) and a Bayesian approach using Markov Chain Monte Carlo to investigate the posterior distribution of the model parameters. The Bayesian approach is useful in examining parameter covariances, model selection, and variability in the resulting scattering length density profiles.
The package is designed to facilitate reproducible research; its use in \Jupyter\ notebooks, and subsequent distribution of those notebooks as supporting information, permits straightforward reproduction of analyses.
\end{abstract}

@@

The \Objective\ class uses the \ReflectModel\ and a dataset, \DataD, to calculate $\chi^2$, log-likelihood (Equation~\ref{eqn:2}), log-prior, residuals, and the generative model.
The \DataD\ object has \emph{x, y, y\_err, x\_err} attributes to represent $Q$, $R$, $\mathrm{d}R$, $\mathrm{d}Q$. The \DataD\ object reads a three or four column plain-text datafile. A three column dataset represents $Q$ (\si{\per\angstrom}), $R$, $\mathrm{d}R$ (1 standard deviation). A four column dataset represents $Q$, $R$, $\mathrm{d}R$, $\mathrm{d}Q$ (\si{\per\angstrom}). Extending \DataD\ would allow other formats to be read. One example of this could be an wavelength dispersive file using $\Omega$, $\lambda$ data instead of $Q$; such as that used in energy scanned X-ray reflectometry, or sometimes produced by wavelength dispersive neutron reflectometers. In such a case \ReflectModel\ could be subclassed to make full use of this energy dispersive information.
An \Objective\ can be given a \Transform\ object to permit fitting as $\log_{10} R$ vs $Q$, $RQ^4$ vs $Q$; the default (no \Transform) is $R$ vs $Q$. Several \Objective\ can be combined to form a \GlobalObjective\ for co-refinement. The object-oriented nature allows \Parameter, \Component\ to be reused in in different \Objective, thereby facilitating co-refinement. For a comprehensive demonstration of multiple contrast co-refinement, see the annotated notebook in the supporting information.

The \Objective\ statistics are used directly by the \CurveFitter\ class to perform least-square fitting with the functionality provided by the \SciPy\ package (DifferentialEvolution, Levenberg--Marquardt, LBFGSB - Limited Broyden--Fletcher--Goldfarb--Shanno with bounds).
\CurveFitter\ can also perform Bayesian Markov Chain Monte Carlo (MCMC) sampling of the system, examining the posterior probability distribution of the parameters, Equation~\ref{eqn:1}. The posterior distribution is proportional to the product of the prior probability and the likelihood (or the sum of the log-probabilities):
%
\begin{gather} 
\label{eqn:1}\ p(\theta | D, I) \propto p(\theta | I)\times p(D | \theta, I)\\
p(D | \theta, I) = -\frac{1}{2} \sum \left[\left(\frac{y_n - model} {\sigma_n}\right)^2 + \log(\sigma_n^2)\right]\label{eqn:2}
\end{gather}
%
The prior, $p(\theta | I)$, is the probability distribution function for a parameter, $\theta$, given preexisting knowledge of the system, $I$, as outlined above.
The likelihood (Equation~\ref{eqn:2}), $p(D | \theta, I)$, is the probability of the observed data, $D$, given the model parameters and other prior information. It is calculated from the measured data, $y_n$ (with uncertainties $\sigma_n$), and the generative model. The likelihoods used in MCMC and least-squares assume that the measurement errors are normally distributed, Equation~\ref{eqn:2}. However, other types of measurement errors (e.g. Poissonian) could be implemented by a subclass of \Objective\ overriding the log-likelihood method.

The posterior probability, $p(\theta | I, D)$, is examined by encoding the likelihood and prior distributions and using the \emcee\ and \ptemcee\ packages to perform affine invariant ensemble sampling \cite{emcee, ptemcee}. \ptemcee\ is a fork of \emcee\ that implements parallel tempering for characterisation of multi-modal probability distributions; different modes can be traversed by chain populations at higher `temperatures', while individual modes are efficiently explored by chains at lower `temperatures'.

At the end of an MCMC run the parameter set possesses a number of samples (called a `chain'), whose statistics reveal the distribution and covariance of the parameters, the spread of the model around the data, and in a reflectometry context the range of structures that are consistent with the data.
The chain statistics are used to update each \Parameter\ value, and assign a standard error. For the sampling, these represent the median and half the $[15.87, 84.13]$ percentile range respectively; the latter approximating the standard deviation for a normally distributed statistic.

Parallelisation of the sampling is automatic, and can use MPI on a cluster for greater parallelisation. Visualisation of the samples produced by MCMC sampling is conveniently performed using the \corner\ package for scatter plot matrices \cite{corner}.

A browser-based graphical user interface (GUI) \cite{ipywidgets} is available for fitting within a \Jupyter\ notebook, Figure~\ref{fig:gui}. This GUI uses slab based models for analysing a single dataset, and can be used as a stepping point for building more advanced models via the `To code' code generation button. A fully functional web-based reflectometry analysis notebook is currently available \cite{Nelson2018}.

@@

\subsection{Example data analysis}

@@

The parameters used in the \LipidLeaflet\ component are: area per molecule ($A$), thicknesses for each of the head and tail regions ($t_x$), sums of scattering lengths of the head and tail regions ($b_x$), partial volumes of the head and tail groups ($V_x$), roughness between head and tail region, and SLDs of the solvents for the head and tail group ($\rho_{x,\mathrm{solv}}$). The overall SLD of each of the head and tail group regions are given by:

\begin{gather} 
\label{eqn:3} \phi_{x} = \frac{V_x}{At_x}\\
\rho_x =  \phi_{x} \frac{b_x}{V_x} + (1 - \phi_{x})\rho_{x,\mathrm{solv}} \label{eqn:4}
\end{gather}
The approach used in \LipidLeaflet\ component ensures that there is a 1:1 correspondence of heads to tails.
By default the head and tail solvents are assumed to be the same as the solvent that is used throughout the \Structure. This will be the case when using \LipidLeaflet\ for a solid-liquid reflectometry experiment. However, at the air-liquid, or liquid-liquid interfaces the solvent for the head and tail region may be different, and it is possible to use different solvent SLDs for each. We note that the \LipidLeaflet\ component may also be used to describe other amphiphiles adsorbing at an interface.

@@

The \Jupyter\ notebook used for the analysis, \emph{lipid.ipynb}, is available in the supporting information. The corner plot (Figure~\ref{fig:corner}) produced from the MCMC analysis shows the covariance between parameters, with an area per molecule of \SI{57.0 \pm0.15}{\square\angstrom}. Figure~\ref{fig:global_fit} shows the probability distribution of the generative model around the data and in the SLD profile. These are obtained by plotting a subset of samples from the MCMC chain. The spread in SLD profiles is used to determine what range of structures is consistent with the data. Multi-modalities in these SLD profiles can be due to statistical uncertainties, the $Q$ ranges measured, and the loss of phase information in NR \cite{Majkrzak1999, Heinrich2009}.

\begin{figure}%
\centering
\label{fig:global_fit}%
\includegraphics[width=85mm]{./supporting_information/global_fit.pdf}%
%TODO: series labels on plots; delete points?;
%TODO: artefact in high q d2o fit; truncate at Q = 0/35 ?
%TODO: h2o fit isn't great about 0.20

@@

\begin{figure}
  \includegraphics[width=175mm]{./supporting_information/corner.pdf}
  \caption{Corner plot for the varying parameters of DMPC bilayers measured at three contrasts. The sampling took $\sim$13 minutes on a \SI{3.2}{GHz} quad-core computer for 30 saved steps, corresponding to 6000 samples, with the steps being thinned by a factor of 100. A larger scale image is available in supporting information.}
  \label{fig:corner}
\end{figure}

@@

Each submodule in \refnx\ possesses its own unit testing code for checking that the functions and classes in the module operate correctly, both individually and collectively. For example, there are tests that check that the reflectivity of a model is calculated correctly, or that the behaviour of a function is correct for the different possible inputs and code paths through it. Since the test suite is an integral part of the package each installation is testable. In addition, there is a benchmarking suite to track changes in performance, specifically the speed of critical calculations, over time.

The source code for \refnx\ is held in a version controlled git repository hosted on GitHub.\footnote{https://www.github.com/refnx/refnx} The standard GitHub workflow is followed in which contributors create their own `fork' of the main \refnx\ repository, and create a feature branch to which they make modifications. They then submit a pull request (PR) against the main repository. The modifications made in the PR are checked on continuous integration (CI) web-services that run the test suite against a matrix of Python versions on the macOS, Linux and Windows operating systems. Features are merged into the main repository if all tests pass, and if manual code review concludes that the changes are scientifically correct, of sufficiently high standard, and useful. When a sufficient number of features have accumulated, a new release is made. Successive releases have an incrementing semantic version number which can be obtained from the installed package, with each release being given its own Digital Object Identifier (DOI).

The recommended way of using \refnx\ is from a \conda\ environment, which offers package, dependency and environment management \cite{conda}, using the pre-compiled distributables on the \refnx\ conda-forge channel. These distributables are made as part of the release process using the same CI web-services used to test the code. The matrix of distributables covers the major Python versions currently in use, across the macOS, Windows, and Linux operating systems. Alternatively the package can be installed from source, either directly from the git repository, or via \pip\ from the version uploaded to PyPI.\footnote{https://pypi.python.org/pypi/refnx; the installation command is 'pip install refnx'} Building from source requires a C compiler and the \Cython\ and \NumPy\ packages to be installed; further dependencies should be installed to run the test suite to verify that compilation and installation was successful.

\refnx\ is released under the BSD permissive open source licence. In addition its dependencies are released under open source licences which means that use is free of cost to the end user and also the user is free to modify and improve this software.

@@

In order for a given scattering analysis to be fully reproducible by others, a general set of conditions need to be met:
\begin{itemize}
  \item the processed datasets used in the analysis need to be deposited with a journal article, or be freely available. Ideally the raw datasets, and the means to create the processed datasets should also be made available.
  \item the exact software environment needs to be recreatable.
  %FIXME refnx does not address this further down the stack
  %I think this has been addressed.
  \item the exact ordered set of steps taken during the analysis needs to be listed.
\end{itemize} 
Each of these points are often omitted in the literature. For example, the use of different software versions may change the output of an analysis, or the use of a GUI program may preclude recording the full set of steps, or options, applied by a user. Whilst it is unable to meet the first criterion by itself, the use of \refnx\ in a \Jupyter\ notebook can fulfil the other two requirements, providing a little care is taken. As we have already noted, the ordered set of steps to perform the analysis is the \Jupyter\ notebook in which the analysis was performed and this is an artefact able to be archived.

@@

\refnx\ is a powerful tool for least-squares or Bayesian analysis of neutron and X-ray reflectometry data that is ideally usable for reproducible research with \Jupyter\ notebooks, and has been built with extensibility in mind. Its features include: MCMC sampling of posterior distribution for parameters, structural models constructed from modular components with physically relevant parameterisation, algebraic inter-parameter constraints, mixed area models, co-refinement of multiple datasets, probability distributions for parameter bounds used directly for log-prior terms, and a (\Jupyter\ ) \ipywidgets\ GUI.

>>>>

\hyphenation{Lipid-Leaflet}

@@

\begin{abstract}
\refnx\ is a model-based neutron and X-ray reflectometry data analysis package written in Python. It is cross platform, and has been tested on Linux, macOS, and Windows. Its graphical user interface is browser-based, through a \Jupyter\ notebook.
Model construction is modular, being composed from a series of components that each describe a subset of the interface, parameterised in terms of physically relevant parameters (volume fraction of a polymer, lipid area per molecule, etc). The model and data are used to create an objective, which is used to calculate residuals, log-likelihood, and log-prior probabilities of the system. Objectives are combined to perform co-refinement of multiple datasets, and mixed-area models. Prior knowledge of parameter values is encoded as probability distribution functions or bounds on all parameters in the system. Additional prior probability terms can be defined for sets of components, over and above those available from the parameters alone. Algebraic parameter constraints are available.
A choice of fitting approaches is available, including least-squares (global and gradient-based optimizers) and a Bayesian approach using Markov Chain Monte Carlo to investigate the posterior distribution of the model parameters. The Bayesian approach is useful in examining parameter covariances, model selection, and variability in the resulting scattering length density profiles.
The package is designed to facilitate reproducible research; its use in \Jupyter\ notebooks, and subsequent distribution of those notebooks as supporting information, permits straightforward reproduction of analyses.
\end{abstract}

@@

\subsection{Comparison with experimental data}

The \Objective\ class is the comparator of the predicted and measured reflectivities, using the \ReflectModel\ and a dataset, \DataD, to calculate $\chi^2$, log-likelihood (Equation~\ref{eqn:2}), log-prior, residuals, and the generative model.
The \DataD\ object has \emph{x, y, y\_err, x\_err} attributes to represent $Q$, $R$, $\mathrm{d}R$, $\mathrm{d}Q$. As is standard for many reflectometry data files, the \DataD\ object reads a three or four column plain-text datafile. A three column dataset represents $Q$ (\si{\per\angstrom}), $R$, $\mathrm{d}R$ (1 standard deviation). A four column dataset represents $Q$, $R$, $\mathrm{d}R$, $\mathrm{d}Q$ (\si{\per\angstrom}). Extending \DataD\ would allow other formats to be read. One example of this could be a wavelength dispersive file using ($\Omega$, $\lambda$)-data instead of $Q$, such as that used in energy scanned X-ray reflectometry, or sometimes produced by wavelength dispersive neutron reflectometers. In such a case \ReflectModel\ could be subclassed to make full use of this energy dispersive information.

An \Objective\ can be given a \Transform\ object to permit fitting as $\log_{10} R$ vs $Q$, $RQ^4$ vs $Q$; the default (no \Transform) is $R$ vs $Q$. Several \Objective\ objects can be combined to form a \GlobalObjective\ for co-refinement. The object-oriented nature allows \Parameter, \Component\ to be reused in different \Objective\ objects, and this is the basis for linking parameters between samples for co-refinement. For a comprehensive demonstration of multiple contrast co-refinement, see the annotated notebook in the supporting information.

\subsection{Statistical comparison and refinement}

The \Objective\ statistics are used directly by the \CurveFitter\ class to perform least-square fitting with the functionality provided by the \SciPy\ package (DifferentialEvolution, Levenberg--Marquardt, LBFGSB - Limited Broyden--Fletcher--Goldfarb--Shanno with bounds).
\CurveFitter\ can also perform Bayesian Markov Chain Monte Carlo (MCMC) sampling of the system, examining the posterior probability distribution of the parameters, Equation~\ref{eqn:1}. The posterior distribution is proportional to the product of the prior probability and the likelihood (or the sum of the log-probabilities):
%
\begin{gather} 
\label{eqn:1}\ p(\theta | D, I) \propto p(\theta | I)\times p(D | \theta, I)\\
p(D | \theta, I) = -\frac{1}{2} \sum_n \left[\left(\frac{y_n - y_{\mathrm{model},n}} {\sigma_n}\right)^2 + \log(\sigma_n^2)\right]\label{eqn:2}
\end{gather}
%
The prior, $p(\theta | I)$, is the probability distribution function for a parameter, $\theta$, given preexisting knowledge of the system, $I$, as outlined above.
The likelihood (Equation~\ref{eqn:2}), $p(D | \theta, I)$, is the probability of the observed data, $D$, given the model parameters and other prior information. It is calculated from the measured data, $y_n$ (with uncertainties $\sigma_n$), and the generative model, $y_{\mathrm{model},n}$. The likelihoods used in MCMC and least-squares assume that the measurement errors are normally distributed, Equation~\ref{eqn:2}. However, other types of measurement errors (e.g. Poissonian) could be implemented by a subclass of \Objective\ overriding the log-likelihood method.

Confidence in the the results of the fitting come by examining the posterior probability, $p(\theta | I, D)$. In the simplest form, this is a confidence interval for the fitted parameter, but in the case of correlated parameters, a simple confidence interval can be rather misleading. The posterior probability is examined by encoding the likelihood and prior distributions and then using an MCMC algorithm (via the \emcee\ and \ptemcee\ packages) to perform affine invariant ensemble sampling \cite{emcee, ptemcee}
At the end of an MCMC run, the parameter set possesses a number of samples (called a `chain'), each of which are similarly plausible fits.
The statistics of the chain reveals the distribution and covariance of the parameters, the spread of the model-predicted measurements around the data, and in a reflectometry context, the range of structures that are consistent with the data.
The analysis of the chain thus yields a family of plausible SLD profiles and a family of plausible parameter ranges.
The chain statistics are used to update each \Parameter\ value, and assign a standard error. For the sampling, these represent the median and half the $[15.87, 84.13]$ percentile range respectively; the latter approximating the standard deviation for a normally distributed statistic.

The \ptemcee\ package is a variant (a `fork' in open-source software development terms) of the \emcee\ package that has been extended to implement the parallel tempering algorithm for characterisation of multi-modal probability distributions; different modes can be traversed by chain populations at higher `temperatures', while individual modes are efficiently explored by chains at lower `temperatures' \cite{ptemcee}. While \ptemcee\ is not strictly a global fitter, in practice, having multiple populations in the parallel tempering algorithm allows the fitter to converge to the global minimum and escape local minima.

Parallelisation of the sampling is automatic, making full use of multiprocessor machines and can use MPI on a cluster for yet greater parallelisation such as for corefinement of very large data sets. Visualisation of the samples produced by MCMC sampling is conveniently performed using the \corner\ package for scatter plot matrices \cite{corner}, which gives a representation of the probability distribution function for each individual parameter and also the correlation for each pair of parameters. As will be seen later, the plots for two normally distribute and uncorrelated parameters is isotropic, while correlated parameters show significant anisotropy. An evaluation of the impact of hard bounds can also be made by looking for correlation plots where the bounds are clearly truncating the distribution function, allowing the bounds to be re-evaluated and adjusted if necessary.

\subsection{User interface}

A significant motivation in the development of \refnx\ has been the facilitation of reproducible analysis by helping the user describe \emph{how} the analysis was performed. A few lines of computer code is an incredibly powerful description, conveying the details with precision that is hard to match in written text and also being incredibly concise. Analysis code can be constructed in \refnx\ in relatively few lines of code (as illustrated in the supporting information), with example analyses often being sufficient to complete the task. These few lines of Python code can be further extended to produce publication quality plots saved and ready to import into the next publication.

While Python is a popular language for instruction and for data analysis, meaning that the relatively few lines of code required to complete a \refnx\ analysis of a set of experiments is not a huge hurdle, a simpler graphical user interface (GUI) is also provided.
The browser-based GUI is available for fitting within a \Jupyter\ notebook, Figure~\ref{fig:gui}, leveraging the ipywidgets modules \cite{ipywidgets}. The GUI has a `To code' button that turns the current model into the few lines of code required to perform the analysis without using the GUI, thus providing the desired instructions for the reproducible analysis. The ability to generate analysis code allows also makes it a stepping point for building more advanced models independently.
The current GUI is able to create slab based models for analysing a single dataset; a fully functional web-based reflectometry analysis notebook is currently available \cite{Nelson2018}.

@@

\section{Example data analysis with a lipid bilayer}

@@

The parameters used in the \LipidLeaflet\ component are: area per molecule ($A$), thicknesses for each of the head and tail regions ($t_x$), sums of scattering lengths of the head and tail regions ($b_x$), partial volumes of the head and tail groups ($V_x$), roughness between head and tail region, and SLDs of the solvents for the head and tail group ($\rho_{x,\mathrm{solv}}$). The overall SLD of each of the head and tail group regions are given by:
\begin{gather} 
\label{eqn:3} \phi_{x} = \frac{V_x}{At_x}\\
\rho_x =  \phi_{x} \frac{b_x}{V_x} + (1 - \phi_{x})\rho_{x,\mathrm{solv}} \label{eqn:4}
\end{gather}
The approach used in \LipidLeaflet\ component ensures that there is a 1:1 correspondence of heads to tails.
By default the head and tail solvents are assumed to be the same as the solvent that is used throughout the \Structure. This will be the case when using \LipidLeaflet\ for a solid-liquid reflectometry experiment. However, at the air-liquid, or liquid-liquid interfaces the solvent for the head and tail region may be different, and it is possible to use different solvent SLDs for each. We note that the \LipidLeaflet\ component may also be used to describe other amphiphiles adsorbing at an interface.

@@

The \Jupyter\ notebook used for the analysis, \emph{lipid.ipynb}, is available in the supporting information. The corner plot (Figure~\ref{fig:corner}) produced from the MCMC analysis shows the covariance between parameters, with an area per molecule of \SI{57.0 \pm0.15}{\square\angstrom}. Figure~\ref{fig:global_fit} shows the probability distribution of the generative model around the data and in the SLD profile. These families of plausible fits that are obtained by plotting a subset of samples from the MCMC chain. The spread in SLD profiles is used to determine what range of structures is consistent with the data. Multi-modalities in these SLD profiles can be due to statistical uncertainties, the $Q$ ranges measured, and the loss of phase information in NR \cite{Majkrzak1999, Heinrich2009}.

\begin{figure}
\centering
\label{fig:global_fit}%
\includegraphics[width=85mm]{./supporting_information/global_fit.pdf}%
%TODO: series labels on plots; delete points?;
%TODO: artefact in high q d2o fit; truncate at Q = 0/35 ?
%TODO: h2o fit isn't great about 0.20

@@

\begin{figure}
  \includegraphics[width=175mm]{./supporting_information/corner.pdf}
  \caption{Corner plot for the varying parameters of DMPC bilayers measured at three contrasts. The sampling took $\sim$\SI{13}{\minute} on a \SI{3.2}{GHz} quad-core computer for 30 saved steps, corresponding to 6000 samples, with the steps being thinned by a factor of 100. A larger scale image is available in supporting information.}
  \label{fig:corner}
\end{figure}

@@

Each submodule in \refnx\ possesses its own unit testing code for checking that the functions and classes in the module operate correctly, both individually and collectively. For example, there are tests that check that the reflectivity of a model is calculated correctly, or that the behaviour of a function is correct for the different possible inputs and code paths through it. Since the test suite is an integral part of the package each installation is testable. In addition, there is a benchmarking suite to track changes in performance, specifically the speed of critical calculations, over time. This development approach is important to providing assurances to the community that the code is tested and works.

The source code for \refnx\ is held in a publicly accessible version controlled git repository.\footnote{https://www.github.com/refnx/refnx} The standard GitHub workflow is followed in which contributors create their own `fork' of the main \refnx\ repository, and create a feature branch to which they make modifications. They then submit a pull request (PR) against the main repository. The modifications made in the PR are checked on continuous integration (CI) web-services that run the test suite against a matrix of Python versions on the macOS, Linux and Windows operating systems. Features are merged into the main repository if all tests pass, and if manual code review concludes that the changes are scientifically correct, of sufficiently high standard, and useful. When a sufficient number of features have accumulated, a new release is made. Successive releases have an incrementing semantic version number which can be obtained from the installed package, with each release being given its own Digital Object Identifier (DOI).

The recommended way of using \refnx\ is from a \conda\ environment, which offers package, dependency and environment management \cite{conda}, using the pre-compiled distributables on the \refnx\ conda-forge channel. These distributables are made as part of the release process using the same CI web-services used to test the code. The matrix of distributables covers the major Python versions currently in use, across the macOS, Windows, and Linux operating systems. Alternatively the package can be installed from source, either directly from the git repository, or via \pip\ from the version uploaded to PyPI.\footnote{https://pypi.python.org/pypi/refnx; the installation command is `\texttt{pip install refnx}'} Building from source requires a C compiler and the \Cython\ and \NumPy\ packages to be installed; further dependencies should be installed to run the test suite to verify that compilation and installation was successful.

\refnx\ is released under the BSD permissive open source licence. In addition, all of the  dependencies of \refnx\ are released under open source licences which means that use is free of cost to the end user and, more importantly, the user user is free to modify, improve, and inspect this software.

@@

In order for a given scattering analysis to be fully reproducible by others, a general set of conditions need to be met:
\begin{itemize}
  \item the processed datasets used in the analysis need to be deposited with a journal article, or be freely available. Ideally the raw datasets, and the means to create the processed datasets should also be made available.
  \item the exact software environment needs to be recreatable.
  \item the exact ordered set of steps taken during the analysis needs to be listed.
\end{itemize} 
Each of these points are often omitted in the literature. For example, the use of different software versions may change the output of an analysis, or the use of a GUI program may preclude recording the full set of steps, or options, applied by a user. Whilst it is unable to meet the first criterion by itself, the use of \refnx\ in a \Jupyter\ notebook can fulfil the other two requirements, providing a little care is taken. As we have already noted, the ordered set of steps to perform the analysis is the \Jupyter\ notebook in which the analysis was performed and this is an artefact able to be archived.

@@

\refnx\ is a powerful tool for least-squares or Bayesian analysis of neutron and X-ray reflectometry data that is ideally usable for reproducible research with \Jupyter\ notebooks, and has been built with extensibility in mind. Its features include: MCMC sampling of posterior distribution for parameters, structural models constructed from modular components with physically relevant parameterisation, algebraic inter-parameter constraints, mixed area models, co-refinement of multiple datasets, probability distributions for parameter bounds used directly for log-prior terms, and a (\Jupyter) \ipywidgets\ GUI.
